---
title: "Exercise 8 - PCA for data QC"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

For another example of where PCA can be helpful, we will use the `transcriptome.csv` file in the data
folder. This data contains the gene expression values for 5380 genes across 24 samples of a heat
stress experiment (8 timepoints and three replicates). To get this data ready, we will need to read
it in, remove the non-numerical sample names from the data and scale it.

```{r expression_data}
heat_stress <- read_csv("data/transcriptome.csv") %>% 
  column_to_rownames("sample") %>% 
  scale()

# Dataset is 24 * 5380. Show just a small part as an example
heat_stress[1:5, 1:5]
```

This dataset has 24 rows (one for each sample), and 5380 columns (one for each gene). This very wide
dataset is similar in structure to the spectral data shown in the lessons. We would like to perform
PCA on these samples to identify samples with similar expression profiles. This provides a good first
check that our experiment has been succesful before we begin a more detailed analysis.

```{r expression_pca}
# Peform the PCA on the already scaled heat_stress data
heat_stress_pca <- _____

# Check the summary
summary(heat_stress_pca)
```

Now, we would like to plot the samples along the principal components. Since the sample names are of
the form Replicate_Timepoint, we will first separate those variables out using `separate`. To begin,
modify the code block below to show a plot along the first two principal components:
```{r expression_plot}
heat_stress_pca$x %>% 
  as_tibble(rownames = "sample") %>% 
  separate(sample, into = c("replicate","timepoint"), sep = "_") %>% 
  ggplot(aes(_____)) +
  geom_point()
```

Once you have got the basic plot working, try colouring the points by the replicate or the timepoint.

Which of these factors is more associated with the variation in the expression data and what does
that tell you about the experiment?

Sometimes moving beyond the first two PCs can provide valuable information as well. Plot the samples
along PC3 and PC4 and colour them by timepoint. Can you guess how long the heat stress lasted?

(If you struggle distinguishing the different colours, you could try using `geom_text` instead and 
provide the timepoint to the `label` aesthetic.)