---
title: "Exercise 3 - Testing means"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

>Apply a t-test to compare means between two samples.

To ask if the means between two samples are the same, you can use the `t.test()` function. This uses
the same formula notation as the `boxplot()` function did earlier. Remember that the structure of this formula 
is the response variable (what you are measuring) on the left, the grouping variable on the right.

Since the InsectSprays dataset has several groups, you will need to filter out just two to perform 
the t test otherwise you will get an error. Pick a few pairs of sprays to compare (from your boxplots
or summary statistics). 

```{r t_tests}
# sleep t test
sleep_t <- t.test(_____ ~ _____, data = sleep)

sleep_t

# InsectSprays t test for sprays A and B
spray_AB <- filter(InsectSprays, spray %in% c("A","B"))

spray_AB_t <- t.test(_____)

spray_AB_t

# Choose your own pair of sprays to compare:

```

What information does the t test show you, and how do you interpret these result?


## t test summary

A t test is used to ask if the means of two samples (`A` and `B` in the diagram below) are different. 
The test first summarises the data points (the `.`s) into two summary statistics, the means of the 
two groups (`o`) and the amount of variation from those means (`=====`).

The difference between the means (`-------`) is then compared to the variation around the means (`=====`)
to produce a t statistic. A large t statistic means the difference between the means is very large
compared to the variation around each mean.

This t statistic is used to calculate a p value, which tells us the probability of observing a 
t statistic at least this large if we assume that there is no true difference in means between the
two groups (the null hypothesis).

```
A    .   .  ..  .  ... ..  ..  .  ..  .
                =====o=====
                     ------------------------------
                                            ======o======
B                .         .       .   .  ..  .  ... ..  ..  .  ..  ..
__________________________________________________________________________
0          10          20          30          40          50          60
```