---
title: "Exercise 5 - Multi-linear regression"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)

source("stats_helpers.R")
```

For this exercise, we will be using data around how 
[US politicians refer to the President on twitter](https://www.jtimm.net/2019/08/20/referring-to-potus/).
The purpose of this data was to 

> consider how a House Rep’s stance towards (or opinion of) 45 influences the choice of referring 
expression, as well as how this stance aligns with the popularity of 45 in a House Rep’s congressional district.

In other words, do politicians whose electorates voted strongly for Donald Trump tend to refer more
to him using more formal language and vice versa?

### Discussion 

To get started, we will consider a single linear relationship in the data. Visit https://spearce.me/shiny/models/
to view a simplied version of the data from this study. Under the "Draw here" heading, use your mouse
to draw your best guess of what the linear relationship between the data is. Once you are happy with
it, you can right click the combined image under the "Save plot here" heading and copy the image to
post in Slack for discussion.

Did everyone draw a similar relationship? 

The raw data for this plot has been simplified from the original post, and can be downloaded from
the site itself or found in the `data` folder.

Check the data below and see if there are any other variables recorded that you think might influence
the Degree of Formality (`DoF`) for each representative:

```{r check_data}
tweet_data <- read_csv("data/tweet_data.csv")

tweet_data
```

### More Discussion
Additional variables can be added to the plot by selecting the "Add more data variables" checkbox.
Does this new information change your view of the relationships in the data? Draw your new guess at
the relationship and post it in Slack for comparison.


There are four different ways you might go about modelling this relationship. They are summarised 
graphically in the `stats_helpers.R` file and can be seen in the `tweet_models` variable. Each model
facet is labelled with the formula provided to the `lm` function to create it.
```{r all_models, fig.height=6}
tweet_models
```

We will leave discussion of which is the most appropriate model for later, and will instead look at
how to fit and interpret the additive model (`DoF ~ trump_margin + party`), and leave the interactive
model ((`DoF ~ trump_margin * party`)) as an exercise.

To fit a linear model using the Trump vote margin, as well as the representative's party, we include
both terms in the formula we provide to `lm`:

```{r fit_additive}
additive_model <- lm(DoF ~ trump_margin + party, data = tweet_data)

# Check results
summary(additive_model)
```

How do we interpret this result? Remember that the formula for the multivariate model is:
                $$Y = B_1*X_1 + B_2*X_2 + \dots + B_0 + \epsilon$$
In our example, $Y$ is `DoF`, $X_1$ is `trump_margin`, and $X_2$ is the `party`.

Running `lm` determines the 'best' values for the $B$ variables, and these values are what is 
displayed in the summary table. $B_0$ is the (Intercept) coefficient and is the hypothetical value for
the Degree of Formality for a Representative where all other variables are zero.

$B_1$ is the `trump_margin` coefficient, and tells us how much DoF increases for each unit increase 
in `trump_margin`. So an increase in the value of `trump_margin` by 1 is associated with an increase
in the Degree of Formality measure by 0.022.

$B_2$ is a little different because `party` is not a numerical variable like `trump_margin` is. 
Instead, $B_2$ is equal to zero if the party is Democrat (the
first alphabetically, and so the 'default' level of the factor), and equal to the `partyRepublican` 
coefficient when the party of the representative is Republican. So for any given value of `trump_margin`, 
a Republican is expected to have a Degree of Formality measure 2.96 higher than a Democrat.

Compare the values with the fitted lines in the plot above. Do these values make sense?

Finally, don't forget to check your model to check that your assumptions around the residuals hold.

```{r additive_check}
resid_panel(additive_model)
```


#### Trying it out

Now, give this a try with the interactive model (`Dof ~ trump_margin * party`)
```{r fit_interactive}
interactive_model <- _____
```

Then, check the results with `summary`, and assess the model residuals.
```{r diamonds_model}
# Check model results
summary(interactive_model)

# Check model assumptions
resid_panel(interactive_model)
```
Can you write out a paragraph explaining the results of this interactive model? Start with the 
coefficient terms that are the same as in the additive model. Then try to work out what the new 
coefficient represents. (**Hint:** Compare the slopes of the fitted lines in the demonstration plot).