---
title: "Exercise 7 - PCA and regression"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Principal Component Analysis

To introduce PCA, we will work with the inbuilt `mtcars` dataset. First, we will select the variables 
and scale them all to have mean equal to zero and standard deviation equal to one.

```{r mtcars_data}
# Removing am and vs because they are categorical variables
# Removing mpg because we will use it later to show how we can use PCA
mtcars_scaled <- select(mtcars, -mpg, -am, -vs) %>% scale

head(mtcars_scaled)
```

To perform PCA, we will use the `prcomp` function. This function takes a data frame of numeric 
values that provide the data for the PCA. We have already scaled the data, but if you have not
you can provide `prcomp` the argument `scale. = TRUE` to have it do the scaling for you. (note the 
fullstop at the end of scale though).

```{r mtcars_PCA}
mtcars_pca <- prcomp(mtcars_scaled)

summary(mtcars_pca)
```

The summary of the PCA results tells us how much of the variance in our data is explained by each 
component. In this case, the first component (PC1) explains 60% of the variance, and the second (PC2)
explains 26% for a combined total of 86% explained by both PC1 and PC2. You can also see this in 
graphical form (also called a screeplot) by calling `plot` on the PCA result.

```{r mtcars_screeplot}
plot(mtcars_pca)
```


The "rotated" data can be accessed using `mtcars_pca$x` and can be used to plot the data along the 
principal components.

```{r mtcars_plot}
head(mtcars_pca$x)

# This is stored as a matrix. Good for mathematical functions, not so 
# good for tidyverse manipulation and plotting. So we will convert it.
mtcars_pca_plot <- mtcars_pca$x %>% 
  as_tibble(rownames = "car") %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point() 

mtcars_pca_plot
```

The values used to transform the original data into the principal components can be found in 
`mtcars_pca$rotation`.

This can be used to help understand how much each original variable contributes to each principal
component.

```{r mtcars_rotation}
mtcars_pca$rotation
```
For example, the PC1 values are high when the `cyl`, `disp`, `hp`, `wt`, and `carb`
variables are high, and low when the `drat`, `qsec`, and `gear` variables are high.

#### Using PCA in regression

We can use these PCA results in a linear regression analysis. They may have several benefits in this
case since the transformed variables are uncorrelated from each other, and the first few components 
will explain most of the variation in a dataset.

For example, running a linear regression on the original mtcars dataset:
```{r mtcars_lm}
# . is a shortcut for "all other variables"
lm(mpg ~ ., data = mtcars) %>% 
  summary()
```

If instead, we combine the mpg variable with our calculated principal component variables:
```{r mtcars_pca_lm}
#Combine
rotated_mtcars <- as_tibble(mtcars_pca$x)
rotated_mtcars$mpg <- mtcars$mpg

#Model
lm(mpg ~ ., data = rotated_mtcars) %>% 
  summary()
```

And since PC1/PC2/PC3 explained a total of 92% of the variance in the dataset between them, we would
probably only have needed to include them from the start.
