---
title: "Exercise 9 - k-means clustering"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
source("stats_helpers.R")
```

# k-means clustering

The function `kmeans` performs k-means clustering on a dataset. The function takes the dataset you
wish to cluster, and the number of clusters you wish to fit.

We will revisit our PCA on the mtcars dataset and cluster the cars along the first two principal
components. There are several methods to choose the 'best' number of clusters, but for this example
we will arbitrarily choose to cluster into five groups.

```{r mtcars_kmeans}
#Select just PC1/PC2
mtcars_PC1_PC2 <- mtcars_pca$x[,c("PC1", "PC2")]

#Cluster mtcars on PC1/PC2
set.seed(30)
mtcars_cluster <- kmeans(mtcars_PC1_PC2, centers = 5)

mtcars_cluster
```

To extract the clustering assignment for each car, we can use `mtcars_cluster$cluster`. To find the 
coordinates for the final position of the centroid for each cluster, use `mtcars_cluster$centers`.

We can overlay this new information on top of the plot we produced earlier

```{r mtcars_cluster_plot}
# Data conversion
car_clusters <- as.character(mtcars_cluster$cluster)
car_centroids <- as_tibble(mtcars_cluster$centers)

# Plot it
mtcars_pca_plot +
  geom_point(aes(colour = car_clusters)) +
  geom_point(data = car_centroids)
```

### Try it out

Apply k-means clustering to the first two principal components of the gene expression data you used 
above. Overlay this clustering information on a plot of the samples. How many clusters work best 
with the data?